A Process for the Emulation of Comparative Oncology Trials with Real-world Evidence (ENCORE)

### 2025-02-16 manuscript/manuscript.qmd
added ConcertAI comments

However, the validity and transportability of results derived between RWE studies and RCTs can depend on many factors and frequently referenced limitations include missing data, small sample sizes, data discontinuity,[@Merola2022; @joshua2022longitudinal] rapid changes in guideline treatment patterns and the inability to measure and emulate common eligibility criteria, prognostic factors, and standardized response assessments in real-world data (RWD).[@rider2024emulations] While there are already published examples of oncology trial emulations,[@rider2024emulations; @merola2023aetion; @merola2024calibrating] a systematic and scaled approach to emulate a diverse set of different oncology trials in multiple heterogeneous databases is necessary to gain confidence in the validity of RWE studies and to provide context as to which questions can be validly answered.

@@

The rationale and operationalization of each criterion is listed in detail in @tbl-criteria. We will mainly consider pivotal interventional, randomized trials after 2011 because treatment guidelines among included cancer indications have undergone significant changes in recent years. Due to the rapid adoption of new breakthrough therapies in routine care, it is unlikely to find patients who may be still treated with outdated treatment regimens in current clinical practice. We also did not include trials with results published too recently. This decision was in order to allow for enough data and follow-up time accrual in the databases used for this project. Although there have been substantial methodological advancements to increase our understanding on the emulation and comparison of real-world progression-free survival (PFS) and objective response rates (ORR) to a RECISTv1.1[@eisenhauer2009new]-based PFS and ORR assessment in RCTs [@ThanCCR; @mckelvey2024evaluation], imaging-based evaluations still hold a level of granularity which may not be necessarily reflected in chart-abstracted assessments of a patient's progression in routine care.[@rwdRECIST; @rivera2022friends] Given this, and the large number of other methodological challenges like missing data, small sample sizes, data discontinuity and rapidly changing guideline treatments, we focus on the emulation of overall survival (OS) as the endpoint of interest. Therefore, we include only trials that have reported overall survival (OS) as one of the pre-specified endpoints in the protocol.

>>>>

However, the validity and transportability of results derived between RWE studies and RCTs can depend on many factors and frequently referenced limitations include missing data, small sample sizes, data discontinuity,[@Merola2022; @joshua2022longitudinal] adoption of changes in guidelines in real-world care and the inability to measure and emulate common eligibility criteria, prognostic factors, and standardized response assessments in real-world data (RWD).[@rider2024emulations] While there are already published examples of oncology trial emulations,[@rider2024emulations; @merola2023aetion; @merola2024calibrating] a systematic and scaled approach to emulate a diverse set of different oncology trials in multiple heterogeneous databases is necessary to gain confidence in the validity of RWE studies and to provide context as to which questions can be validly answered.

@@

The rationale and operationalization of each criterion is listed in detail in @tbl-criteria. We will mainly consider pivotal interventional, randomized trials after 2011 because treatment guidelines among included cancer indications have undergone significant changes in recent years. Due to the rapid adoption of new breakthrough therapies in routine care, it is unlikely to find patients who may be still treated with outdated treatment regimens in current clinical practice. We also did not include trials with results published too recently. This decision was in order to allow for enough data and follow-up time accrual in the databases used for this project. Although there have been substantial methodological advancements to increase our understanding on the emulation and comparison of real-world progression-free survival (PFS) and objective response rates (ORR) to a RECISTv1.1[@eisenhauer2009new]-based PFS and ORR assessment in RCTs [@ThanCCR; @mckelvey2024evaluation], imaging-based evaluations still hold a level of granularity which may not be necessarily reflected in chart-abstracted assessments of a patient's progression in routine care.[@rwdRECIST; @rivera2022friends] In addition, the timing and cadence of intervals between progression assessments can differ between RCTs and routine care which may result in measurement error and bias.[@ackerman2024measurement] Given this, and the large number of other methodological challenges like missing data, small sample sizes, data discontinuity and rapidly changing guideline treatments, we focus on the emulation of overall survival (OS) as the endpoint of interest. Therefore, we include only trials that have reported overall survival (OS) as one of the pre-specified endpoints in the protocol.
